Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-02-19 draft.tex
formulation changes until the explanations section

		{\normalsize im Studiengang \\
			\Studiengang  im \semester. Semester \\
		}
		\vspace{0.5cm}
		{\normalsize vorgelegt von}
		\vspace{0.5cm}

@@

	mel-spectrograms in combination with convolutional neural networks, achieving well-performing
	results on the ASVSpoof2019 dataset \cite{wang_asvspoof_2020}.

@@

	A further work \cite{yan_initial_2022} makes use of MFCCs, in combination with other
	hand-crafted imperceptible features such as linear frequency cepstral coefficients (LFCCs) to
	not only classify audio samples as synthetic or authentic, but also to correctly classify the
	vocoder used to produce the sample with nearly perfect accuracy on a handcrafted dataset.

@@

	Fourier transforms, such as the short term Fourier transform (STFT) to achieve accurate
	reuslts on the ASVSpoof2017 dataset \cite{delgado_asvspoof_2018}.

@@

	Yang et al. \cite{yang_robust_2024} demonstrate a multi-feature comparison, as well as a

@@

	In-the-Wild \cite{muller_does_2022}, presenting their results for each features, as well as

@@

	The work of Barrington et al. \cite{barrington_single_2023} explores the idea of classification
	using perceptible features, highlighting their potential for improving explainability in the
	space of deepfake audio detection. They do, however also demonstrate a drop in performance in
	comparison with imperceptible hand-crafted and deep-learning features in their experiments,
	which were performed on a combination of synthetic and real audio datasets.

@@

	perceptron-architecture with a single hidden layer, and trained and tested using the dataset

@@

	experiments, which were performed as well on the dataset of the ASV2022 Challenge
	\cite{yi_add_2024}.

@@

	Another example is presented by Haq et al. \cite{haq_multimodal_2023}, who use
	the changes in emotional state as an input feature and represent "unlikely" changes on a graph
	so that it can be understood by an eventual end user. In this case, they achieved their
	results by combining the output of fake video and fake audio classifiers in order to produce a
	final classification for a video sample with audio. Testing against the presidential deepfake
	dataset \cite{sankaranarayanan_presidential_nodate}, they achieve impressive results in
	comparison to the standing benchmark on the dataset at the time.

@@

		The model used for this experiment was made, trained and evaluated using the tools

@@

			understandable to the end user, a focus was placed on using multiple perceptible
			features as inputs for the classifier. Two imperceptible features were also added with
			the hypothesis, based on previous research, that they would positively of increasing
			the model performance.

@@

			The general architecture is designed to allow the combination of diverse features into

@@

			shape. This is achieved by using a separate model for each feature, with its own input

@@

			together, processed though a final model that pools into a final output value (which I

@@

			required in input preparation, but also allows for more flexibility in the

@@

			advantage to the generation of explanations, as only the terminus part of the entire
			must be explained to interpret the importance of each input feature. Further
			clarification is in the following section.

@@

			into the terminus model at the end, which produces the final result.
			Figure \ref{fig:submodel} offers a deeper look into the general architecture of a
			singular sub-model in our method. The sub-models begin with an input layer, taking
			either a vector or matrix depending on the dimensionality of the feature to which the
			model is tailored. The inputs are then passed into alternating convolutional and
			pooling layers, in order to increase the localized pattern detection capability within
			the sub-models. Figure \ref{fig:term_model} is a generalized representation of the

@@

			The black-box model described in the previous section was trained on the first portion

@@

			exact experiment (more details are provided in the next section). Evaluation was then
			performed on the remaining samples in the dataset. Training batches consisted of 100
			samples each with every window position for each sample and an upper limit of

@@

		The explanations using the previously described surrogate model are generated with the
		Local Interpretable Model-agnostic Explanation (LIME) method \cite{ribeiro_why_2016}. This

@@

		for individual samples, as well as how the usefulness of the explanations can be evaluated
		in general.

>>>>

		{\normalsize im Studiengang \\
			\Studiengang \ im \semester. Semester \\
		}
		\vspace{0.5cm}
		{\normalsize vorgelegt von}
		\vspace{0.5cm}

@@

	mel-spectrograms in combination with an architecture based on convolutional neural networks,
	achieving well-performing results on the ASVSpoof2019 dataset \cite{wang_asvspoof_2020}.

@@

	A further work from Yan et al. \cite{yan_initial_2022} makes use of MFCCs, in combination with
	other hand-crafted imperceptible features such as linear frequency cepstral coefficients
	(LFCCs) to not only classify audio samples as synthetic or authentic, but also to correctly
	classify the vocoder used to produce the sample with nearly perfect accuracy on a handcrafted
	dataset.

@@

	Fourier transforms, such as the short term Fourier transform (STFT). The aforementioned
	example uses these features combined with others such as the aforementioned mel-spectrogram
	based features. They also pair these features with a convolutional architecture and achieve
	notably accurate results on the ASVSpoof2017 dataset \cite{delgado_asvspoof_2018}.

@@

	Yang et al. \cite{yang_robust_2024} demonstrate a comparison of multiple features, as well as a

@@

	In-the-Wild \cite{muller_does_2022}, presenting their results for each feature, as well as

@@

	The work of Barrington et al. \cite{barrington_single_2023} explores the idea of
	classification using perceptible features, highlighting their potential for improving
	explainability in the space of deepfake audio detection, although they only implement a
	classifier and not an explainer. They do, however, also demonstrate a drop in performance when
	using perceptible features in comparison with imperceptible hand-crafted and deep-learning
	features in their experiments, which were performed on a combination of synthetic and real
	audio datasets.

@@

	perceptron architecture with a single hidden layer and trained and tested using the dataset

@@

	experiments in comparison to experiments run with only perceptible or imperceptible features,
	which were performed as well on the dataset of the ASV2022 Challenge \cite{yi_add_2024}.

@@

	Another example is presented by Haq et al. \cite{haq_multimodal_2023}, who use the changes in
	emotional state as an input feature and represent "unlikely" changes on a graph so that it can
	be understood by an eventual end user. In this case, they achieved their results by combining
	the output of fake video and fake audio classifiers in order to produce a final classification
	for a video sample with audio. Testing against the presidential deepfake dataset
	\cite{sankaranarayanan_presidential_nodate}, they achieve impressive results in comparison to
	the standing benchmark on the dataset at the time.

@@

		We created, trained and evaluated the model used for our experiments using the tools

@@

			understandable to the end user, we placed a focus on using multiple perceptible
			features as inputs for the classifier. We added used two imperceptible features with
			the hypothesis, based on previous research \cite{barrington_single_2023,
			chaiwongyen_contribution_2022,chaiwongyen_deepfake-speech_2023,li_comparative_2022},
			that they would positively of increasing the model performance.

@@

			We designed our model architecture to allow the combination of diverse features into

@@

			shape. We achieved this by using a separate model for each feature, with its own input

@@

			together and processed though a model that pools into a final output value (which we

@@

			required for the input preparation, but also allows for more flexibility in the

@@

			advantage to the generation of explanations, as only the terminus part of the hybrid
			model must be explained to interpret the importance of each input feature. Further
			clarification is in §\ref{sec:method_exp}.

@@

			into the terminus model at the end, which produces the final result. Figure
			\ref{fig:submodel} offers a deeper look into the general architecture of a singular
			sub-model in our method. The sub-models begin with an input layer, taking either a
			vector or matrix depending on the dimensionality of the feature to which the model is
			tailored. The inputs are then passed into alternating convolutional and pooling
			layers, in order to increase the localized pattern detection capability within the
			sub-models. Figure \ref{fig:term_model} is a generalized representation of the

@@

			We trained fhe black-box model described in the previous section on the first portion

@@

			exact experiment (more details in §\ref{sec:experimental_results}). Evaluation was
			then performed on the remaining samples in the dataset. Training batches consisted of
			100 samples each with every window position for each sample and an upper limit of

@@

		\label{sec:method_exp}
		We generate the explanations using the previously described surrogate model with the Local
		Interpretable Model-agnostic Explanation (LIME) method \cite{ribeiro_why_2016}. This

@@

		for individual samples, as well as how we evaluate the usefulness of the explanations in
		at a global scale.
